A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2021-11-29 neo.Rmd
modifications and additions

```{r tabulate}
tneo <- toTpch(Surv(enter, exit, event) ~  
                   sex + socStatus + illeg + parity + cold, 
               data = neo, cuts = 0:28)
dim(tneo)
head(tneo)
summary(tneo)
saveRDS(tneo, file = "data/tneo.rds")
```

Note how it works: For each combination of *socStatus*, *illeg*, *parity* *cold*
(*cold* is temperature above minimum average), and *age*, the number of deaths (*event*)
and the *exposure* time (in days) are calculated. Note that the tabulated data frame is 
a tenth in size compared to the original file, and it is also *anonymised*.

@@

>>>>

```{r tabulate}
tneo <- toTpch(Surv(enter, exit, event) ~  
                   sex + socStatus + illeg + parity + cold, 
               data = neo, cuts = 0:28)
dim(tneo)
head(tneo)
summary(tneo)
saveRDS(tneo, file = "t_data/tneo.rds")
```

Note how it works: For each combination of *socStatus*, *illeg*, *parity* *cold*
(*cold* is temperature above average minimum), and *age*, the number of deaths (*event*)
and the *exposure* time (in days) are calculated. Note that the tabulated data frame is 
a tenth in size compared to the original file, and it is also *anonymous*.

@@

# Effects of time transformation and tabular data.

As an example of time transformation, we study the *log-cube* transformation used by 
Bourgeouis-Pichat [@bp51a; @bp51b] in his study of endogeneous and exogeneous 
factors explaining infant mortality. It is given by

\begin{equation}
g(t) = \bigl(log(1 + t)\bigr)^3, \quad 0 \le t \le 365,
\end{equation}

where $t$ is measured in days.

So what happens if data are time-transformed before tabulation? Let us see:
First the transformation of the data frame *neo*:

```{r gtrans}
g <- function(t) (log(1 + t))^3
gneo <- neo
gneo$enter <- g(neo$enter)
gneo$exit <- g(neo$exit)
head(gneo)
```

Next tabulation: We need to transform the chosen cut points, if we want comparability, so

```{r tgneo}
tgneo <- toTpch(Surv(enter, exit, event) ~ 
                    sex + socStatus + illeg + parity + cold, 
               data = gneo, cuts = g(0:28))
head(tgneo)
#summary(tgneo)
```

Note the clumsy levels of the factor *age*. Not important. Let us try a "piecewise constant ..."

```{r tpchreg_g, results = 'asis'}
(system.time(fit.pch_g <- tpchreg(oe(event, exposure) ~ 
                                    sex + socStatus + illeg + parity + cold,
                                time = age, data = tgneo)))
dr2_g <- drop1(fit.pch_g, test = "Chisq")
ltx(fit.pch_g, dr = dr2_g, label = "tab:tpchreg_g", 
    caption = "Piecewise constant proportional hazards model with time transformation.")
```

Compare the Tables \@ref(tab:tpchreg) and \@ref(tab:tpchreg_g). Identical except for *TTR*
(Total Time at Risk). And graphs will be stretcehd.

```{r haz_g, fig.cap = "Baseline hazard function, transformed time.",  fig.height = 4}
par(las = 1)
plot(fit.pch_g, fn = "haz")
abline(h = 0)
```

We can spot how the transformation drives the the hazard function to *constant* 
over time, say after $ g(t) = 10$, that is, after around 8 days.

### 2021-11-29 paper.Rmd
modifications and additions

In accordance with the Bourgeouis-Picat ideas, we group the analyses into two parts, 
*neonatal* (first 28 days) och *postneonatal* mortality (days 29--365).

>>>>

In accordance with the Bourgeouis-Pichat ideas, we group the analyses into two parts, 
*neonatal* (first 28 days) och *postneonatal* mortality (days 29--365).
